---
title: "Uber New York"
author: "Subash chander swaminathan"
date: "18/03/2021"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Libraries
```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```
Import CSV files into dataframe
```{r}
apr_data <- read.csv("uber-raw-data-apr14.csv")
may_data <- read.csv("uber-raw-data-may14.csv")
jun_data <- read.csv("uber-raw-data-jun14.csv")
jul_data <- read.csv("uber-raw-data-jul14.csv")
aug_data <- read.csv("uber-raw-data-aug14.csv")
sep_data <- read.csv("uber-raw-data-sep14.csv")
```
created a vector of colors to plot
```{r}
colors = c("#CC1011", "#665555", "#05a399", "#cfcaca", "#f5e840", "#0683c9", "#e075b0")
```
Bind all dataframe together
```{r}
data_2014 <- rbind(apr_data,may_data, jun_data, jul_data, aug_data, sep_data)
```
Convert the format of Date.Time column to data and time
```{r}
data_2014$Date.Time <- as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S")
```

```{r}
data_2014$Time <- format(as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
```

```{r}
data_2014$Date.Time <- ymd_hms(data_2014$Date.Time)
```
Extract the day, month, year and day of the week data to analyse
```{r}
data_2014$day <- factor(day(data_2014$Date.Time))
data_2014$month <- factor(month(data_2014$Date.Time, label = TRUE))
data_2014$year <- factor(year(data_2014$Date.Time))
data_2014$dayofweek <- factor(wday(data_2014$Date.Time, label = TRUE))
```
Extract hour, minutes and seconds also to analyse
```{r}
data_2014$hour <- factor(hour(hms(data_2014$Time)))
data_2014$minute <- factor(minute(hms(data_2014$Time)))
data_2014$second <- factor(second(hms(data_2014$Time)))
```
Trips by hours in a day
```{r}
hour_data <- data_2014 %>% group_by(hour) %>% dplyr::summarize(Total = n())
datatable(hour_data)

ggplot(hour_data, aes(hour, Total)) + 
  geom_bar( stat = "identity", fill = "steelblue", color = "red") +
  ggtitle("Trips Every Hour") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma)
```
Trips by everyday of a month
```{r}
day_group <- data_2014 %>% group_by(day) %>% dplyr::summarize(Total = n()) 

datatable(day_group)

ggplot(day_group, aes(day, Total)) + 
  geom_bar( stat = "identity", fill = "steelblue") +
  ggtitle("Trips Every Day") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma)

day_month_group <- data_2014 %>% group_by(month, day) %>% dplyr::summarize(Total = n())

ggplot(day_month_group, aes(day, Total, fill = month)) + 
  geom_bar( stat = "identity") +
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors)
```
Trips by month in a year
```{r}
month_group <- data_2014 %>% group_by(month) %>% dplyr::summarize(Total = n()) 
datatable(month_group)

ggplot(month_group , aes(month, Total, fill = month)) + 
  geom_bar( stat = "identity") +
  ggtitle("Trips by Month") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors)
```
Trips by day and month
```{r}
month_weekday <- data_2014 %>% group_by(month, dayofweek) %>% dplyr::summarize(Total = n())

ggplot(month_weekday, aes(month, Total, fill = dayofweek)) + 
  geom_bar( stat = "identity", position = "dodge") +
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values = colors)
```
Trips by Base
```{r}
ggplot(data_2014, aes(Base)) + 
  geom_bar(fill = "darkred") +
  scale_y_continuous(labels = comma) +
  ggtitle("Trips by Bases")

ggplot(data_2014, aes(Base, fill = month)) + 
  geom_bar(position = "dodge") +
  scale_y_continuous(labels = comma) +
  ggtitle("Trips by Bases and Month") +
  scale_fill_manual(values = colors)

ggplot(data_2014, aes(Base, fill = dayofweek)) + 
  geom_bar(position = "dodge") +
  scale_y_continuous(labels = comma) +
  ggtitle("Trips by Bases and DayofWeek") +
  scale_fill_manual(values = colors)
```
Heat Map

```{r}
ggplot(month_weekday, aes(dayofweek, month, fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heat Map by Month and Day of Week")

month_base <-  data_2014 %>% group_by(Base, month) %>% dplyr::summarize(Total = n()) 

day0fweek_bases <-  data_2014 %>%
  group_by(Base, dayofweek) %>%
  dplyr::summarize(Total = n()) 

ggplot(month_base, aes(Base, month, fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heat Map by Month and Bases")


ggplot(day0fweek_bases, aes(Base, dayofweek, fill = Total)) +
  geom_tile(color = "white") +
  ggtitle("Heat Map by Bases and Day of Week")

```


